\RequirePackage{ifpdf}
\documentclass[master,tocprelim]{cornell}
%
% tocprelim option must be included to put the roman numeral pages in the
% table of contents
%
% The cornellheadings option will make headings completely consistent with
% guidelines.
%
% This sample document was originally provided by Blake Jacquot, and
% fixed up by Andrew Myers.
%
%Some possible packages to include
\usepackage{graphicx,pstricks}
\usepackage{graphics}
\usepackage{moreverb}
\usepackage{subfigure}
\usepackage{epsfig}
\usepackage{subfigure}
\usepackage{hangcaption}
%\usepackage{txfonts}
\usepackage{palatino}

%new packages added by Yinglei Wang
\usepackage[cmex10]{amsmath}
\usepackage{url}

\def\mywidth{5in}


%if you're having problems with overfull boxes, you may need to increase
%the tolerance to 9999
\tolerance=9999

\bibliographystyle{unsrt}
%\bibliographystyle{IEEEbib}

\renewcommand{\caption}[1]{\singlespacing\hangcaption{#1}\normalspacing}
\renewcommand{\topfraction}{0.85}
\renewcommand{\textfraction}{0.1}
\renewcommand{\floatpagefraction}{0.75}

\title {Flash Memory for Ubiquitous Hardware Security Functions}
\author {Yinglei Wang}
\conferraldate {December}{2013}
\degreefield {M.S.}
\copyrightholder{Yinglei Wang}
\copyrightyear{2013}

\begin{document}

\maketitle
\makecopyright

\begin{abstract}
We demonstrate that unmodified commercial Flash memory can 
provide three important security functions: true random number 
generation, digital fingerprinting and information hiding. Taking advantage of random telegraph 
noise (a type of quantum noise source in highly scaled Flash memory cells) 
enables high quality true random number generation at a rate up to  
10Kbits / second. A scheme based on partial programming exploits process 
variation in threshold voltages to allow quick generation of many unique 
fingerprints that can be used for identification and authentication. 
Information hiding hides 
data within an analog characteristic of Flash, the program time of individual bits. 
Because the technique uses analog behaviors, normal Flash memory operations are
not affected and hidden information is invisible in the data stored in the memory.
Even if an attacker checks a Flash chip's analog characteristics,
experimental results indicate that the hidden information is difficult to distinguish 
from inherent manufacturing variation or normal wear on the device.
Moreover, the hidden data can survive erasure of the Flash memory data.
All
schemes require no change to Flash chips or interfaces, and do not require additional hardware. 
\end{abstract}

\begin{biosketch}
Yinglei Wang attended Peking(Beijing) University in China as an EECS undergraduate student from year 2006 to 2010. 
She entered the MS/PHD program in ECE, Cornell University in Fall, 2010. She worked with Prof. Edwin C. Kan, 
Prof. G. Edward Suh and Prof. Christopher Batten on flash memory security and on-chip network. She will work as
a software engineer at Oracle from January, 2014.
\end{biosketch}

\begin{dedication}
This document is dedicated to my parents and my husband Xuetian Weng, who is now a PHD student in Computer Science.
\end{dedication}

\begin{acknowledgements}
Upon leaving Cornell and Ithaca after spending three years and a half here, I have to admit it is an unexpected ending.
Nevertheless, it is invaluable experience which taught me a lot of things: knowledge, research, relationships and life. 

First, I would like 
to express the deepest appreciation and gratitude to my advisor, Prof. Edwin C. Kan.  He has supported me throughout 
graduate school, and provided the vision and advise necessary to finish my dissertation. More importantly, he gives me the 
freedom to explore new things and tries his best to support me. I would also like to express sincere thanks to my committee member, 
Prof. G. Edward Suh. He has given me so much guidance and help on my research. What's more, he tried to help me in my most 
difficult days, and I really appreciate it. Special thanks to my committee member, Prof. Christopher Batten. 
His computer architecture course is the first computer architecture/organaization course I took and I was deeply attracted to it.
I also learned 
a lot, both technical skills and the methodology of engineering, when working with him. I would also like
to take this opporutnity to express my sincere thanks to one of my course instructors, 
Prof. Jose F. Martinez. In addition to his interesting lectures, He is concerned with my situation and I could feel free to tell him my efforts, worries, concerns and feelings. I really 
appreciate his advice and his attitude towards work and life.

Throughout the years, I have the honor to work with several brilliant graduate and undergradute students. I would like to thank KK Yu, 
Greg Malysa and Shuo Wu for their help and contribution to this topic. I would also like to thank the group memebers, especially 
Kshitij Auluck and Yunfei Ma for their 
inspiring discussions and friendship.

I wish to thank my parents, Deming Wang and Chunhua Wang. Their unconditional love always gives me the warmth and courage to 
carry on. I love them forever. I also hope to thank my husband, Xuetian Weng, who is now a PHD student in Computer Science, Stony Brook 
Unviersity. He can always tolerate my bad temper and complaints. He also encourages me to think about the good things about life 
and be a happy person. I'm so lucky to have his love in my life and I love him. Best wishes to his PHD career.

Sincere thanks and best wishes to all my friends at Cornell.

\end{acknowledgements}

\contentspage
\tablelistpage
\figurelistpage

\normalspacing \setcounter{page}{1} \pagenumbering{arabic}
\pagestyle{cornell} \addtolength{\parskip}{0.5\baselineskip}

\input{introduction.tex}

\input{flash.tex}

\input{rng.tex}

\input{fingerprints.tex}

\input{steg.tex}

\input{conclusion.tex}

\bibliography{sampleThesis}

\end{document}
